Choosing an appropriate scalar transformation to normalise wpd

### 2021-02-07 scalar_transformation.Rmd
document updated

---
title: "Choosing an appropriate scalar transformation to normalise wpd"
author: "Sayani Gupta"
output: pdf_document
---

@@

N01 %>% 
  ggplot(aes(x=log(nx*nfacet), y = (value - intercept)/log(nx*nfacet))) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")  +
  ylab("wpd_norm = (wpd - intercept)/log(nx*nfacet)")

@@

```{r, fig.cap = "The distribution of the transformed wpd is plotted. The shape and the scale of the distribution still changes across different nx and nfacet levels, but seems uniform for higher levels." }
N01 %>% 
  ggplot(aes(x = (value - intercept)/(log(nx*nfacet)))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd") +
  scale_x_continuous(breaks = scales::breaks_extended(3)) +
  xlab("wpd_norm = (wpd - intercept)/(log(nx*nfacet)")

@@

>>>>

---
title: "Choosing an appropriate scalar transformation to normalise wpd"
author: "Sayani Gupta"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    toc: FALSE
---

@@

N01 %>% 
  ggplot(aes(x=log(nx*nfacet), y = (value - intercept)/log(nx*nfacet))) +
  geom_point() + stat_summary(fun=median, geom="line", aes(group=1), color = "blue")  +
  ylab("wpd_norm1 = (wpd - intercept)/log(nx*nfacet)")

@@

```{r, fig.cap = "The distribution of the transformed wpd is plotted. The shape and the scale of the distribution still changes across different nx and nfacet levels, but seems uniform for higher levels." }
N01 %>% 
  ggplot(aes(x = (value - intercept)/(log(nx*nfacet)))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd") +
  scale_x_continuous(breaks = scales::breaks_extended(3)) +
  xlab("wpd_norm1")

@@

# Reduce the heteroskedasticity by the transformation y - a - b*log(n)

In Figure \ref{fig:new-transform}, we see that the transformed wpd value $wpd_{norm2} = wpd - intercept - slope*log(nx*nfacet)$ has lower heteroscedasticity as compared to $wpd_{norm1}$ when plotted across $log(nx*nfacet)$.

```{r new-transform, echo = TRUE, fig.cap = "$wpd_{norm2}$ is plotted against log(nx*nfacet). The relationship is horizontal with reduced heteroscedasticity."}
N01 %>% 
  ggplot(aes(x=log(nx*nfacet), y = (value - intercept - slope*log(nx*nfacet)))) +
  geom_point() + stat_summary(fun=mean, geom="line", aes(group=1), color = "blue") + 
  xlab("wpd_norm2 =  value - intercept - slope*log(nx*nfacet)")
```

In Figure \ref{fig:dist-new-transform}, we see that the distribution of $wpd_{norm2}$ is more similar across nx and nfacets than $wpd_{norm1}$.

```{r dist-new-transform, echo = TRUE, fig.cap = " The distribution of $wpd_{norm2}$ is plotted. The distributions are more similar across different nx and nfacets as compared to $wpd_{norm1}$."}
N01 %>% 
  ggplot(aes(x = ((value - intercept - slope*log(nx*nfacet))))) + 
  geom_density(fill = "blue") +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("wpd_norm2 =  value - intercept - slope*log(nx*nfacet)") +
  scale_x_continuous(breaks = scales::breaks_extended(3))
```

In Figure \ref{fig:dist-new-approx}, we see that the distribution of $wpd_{norm2}$ is similar when normalised with true or approximate estimates of a and b.

```{r dist-new-approx, echo = TRUE, fig.cap = " The distribution of $wpd_{norm2}$ is plotted for approximate estimate of a and b. The distributions are similar when plotted against true estimate of and b."}
N01_compare <- N01 %>% 
  mutate(with_true_est = (value - intercept - slope*log(nx*nfacet)),

                          with_approx_est = (value - 0.04 - 0.003*log(nx*nfacet))) %>% 
  pivot_longer(cols = 5:6,
               names_to = "type_estimate",
               values_to = "value_estimate")

N01_compare %>% 
  ggplot() +
  geom_density(aes(x = value_estimate, fill = type_estimate), alpha = 0.5) +
  facet_grid(nx~nfacet,
             labeller = "label_both") + 
  xlab("value - intercept - slope*log(nx*nfacet)") +
  scale_x_continuous(breaks = scales::breaks_extended(3)) +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#0072B2", "#D55E00")) +
  xlab("wpd_norm2")
```
